\bf The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

### 2021-03-09 paper/forestecology.Rmd
3712 -> 3421 by moving text to figure captions

The output `growth_scbi` is a single data frame of class `sf` that includes variables `growth`, the average annual growth in DBH (cm $\cdot$ y^-1^) for all stems that were alive at both time points, and `geometry`, the `sf` package's encoding of geolocations of type `<POINT>`. In addition the species variable `sp` is returned as a factor.^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set. This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.]

Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using the `geom_sf()` geometry as seen in Figure \ref{fig:scbi-trees} where we plot a random sample of 500 out of the `r nrow(growth_scbi)` trees.

@@

We observe that for this focal tree, we have 4 variables of information on its 20 competitor trees: their unique tree-stem ID number, their distance to the focal tree (all $\leq$ `r comp_dist`), their species, and their basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ where $DBH$ is the value from the earlier of the two censuses in cm. Saving our focal versus competitor information in list-column minimizes redundancy since we do not repeat information on the focal tree 20 times. The spatial distribution of these trees is visualized in Figure \ref{fig:scbi-focal-vs-comp-map}: the dashed circle extends `r comp_dist` m away from the focal tree while all 20 competitor trees are within this circle.

```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4."}
comp_radius <- growth_scbi %>% 
  filter(stemID == 4) %>% 
  st_buffer(dist = comp_dist)

@@

Lastly, the generic for `ggplot2::autoplot()` allows us to plot the posterior distribution of all parameters in Figure \ref{fig:scbi-posterior-viz} (for compactness we include only 3 species).

```{r scbi-posterior-viz, out.width="100%", fig.cap= "Fit model: Posterior distributions of all parameters for three species.", fig.width = 7.5}
# Plot posteriors for only a subset of species
sp_to_plot <- c("litu", "quru", "cagl")

@@

To this end, we use the `foldID` variable defined in Section \ref{spatial-information} whereby all focal trees are assigned to one of `r n_fold` spatially contiguous blocks that act as folds in our cross-validation routine. Figure \ref{fig:scbi-spatial-cross-validation-schematic} presents a schematic illustrating this scheme for fold 1 (bottom-left) as the test set and folds 2--4 as the training set. We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set to compute fitted/predicted values, and compute the RMSE of the observed versus predicted growths. We repeat this procedure 3 more times with each of the three remaining folds acting as the test set and then average all four resulting RMSE's. Furthermore, in order to maintain spatial independence between the test and training set, a "fold buffer" that extend outwards from the boundary of the test set is computed; all trees falling within this fold buffer are excluded from the training set.

```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation: Using the k = 1 fold as the test set, assigning each focal tree to training set, test set, and fold buffer.", echo = FALSE}
fold_number <- 1

@@

The authors thank Ryan Giordano and Jonathan Che for their help with the statistical methodology and Sophie Li for their feedback on package interface. The authors declare no conflicts of interest.

@@

We intend to archive all data and source code for the `forestecology` package as well as this manuscript on GitHub at https://github.com/rudeboybert/forestecology. This repository will be versioned and archived on Zenodo upon acceptance. The 2008 and 2014 Smithsonian Conservation Biology Institute census data loaded in Section \ref{compute-growth}are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and are versioned and archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].

>>>>

The output `growth_scbi` is a single data frame of class `sf` that includes variables `growth`, the average annual growth in DBH (cm $\cdot$ y^-1^) for all stems that were alive at both time points, and `geometry`, the `sf` package's encoding of geolocations of type `<POINT>`. In addition the species variable `sp` is returned as a factor.^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set. This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.] Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using the `geom_sf()` geometry as seen in Figure \ref{fig:scbi-trees}.

@@

We observe that for this focal tree, we have 4 variables of information on its 20 competitor trees: their unique tree-stem ID number, their distance to the focal tree (all $\leq$ `r comp_dist`), their species, and their basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ where $DBH$ is the value from the earlier of the two censuses in cm. Saving our focal versus competitor information in list-column minimizes redundancy since we do not repeat information on the focal tree 20 times. The spatial distribution of these trees is visualized in Figure \ref{fig:scbi-focal-vs-comp-map}.

```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Identify all focal and corresponding competitor trees: The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle."}
comp_radius <- growth_scbi %>% 
  filter(stemID == 4) %>% 
  st_buffer(dist = comp_dist)

@@

Lastly, the generic for `ggplot2::autoplot()` allows us to plot the posterior distribution of all parameters in Figure \ref{fig:scbi-posterior-viz}.

```{r scbi-posterior-viz, out.width="100%", fig.cap= "Fit model: Posterior distributions of all parameters. For compactness we include only three species.", fig.width = 8, fig.height = 6}
# Plot posteriors for only a subset of species
sp_to_plot <- c("litu", "quru", "cagl")

@@

To this end, we use the `foldID` variable defined in Section \ref{spatial-information} whereby all focal trees are assigned to one of `r n_fold` spatially contiguous blocks that act as folds in our cross-validation routine. Figure \ref{fig:scbi-spatial-cross-validation-schematic} presents a schematic illustrating this scheme. We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set to compute fitted/predicted values, and compute the RMSE of the observed versus predicted growths. We repeat this procedure 3 more times with each of the three remaining folds acting as the test set and then average all four resulting RMSE's. Furthermore, in order to maintain spatial independence between the test and training set, a "fold buffer" that extend outwards from the boundary of the test set is computed; all trees falling within this fold buffer are excluded from the training set.

```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation: Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a \"fold buffer.\"", echo = FALSE}
fold_number <- 1

@@

The authors thank Sophie Li for their feedback on the package interface. The authors declare no conflicts of interest.

@@

We intend to archive all data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology. This repository will be archived on Zenodo upon acceptance. The example Smithsonian Conservation Biology Institute census data used are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].

### 2021-03-09 paper/forestecology.tex
3712 -> 3421 by moving text to figure captions

The package designed with ``tidy'' design principles in mind
\citep{wickham_welcome_2019}. Much like many of the \texttt{tidyverse}
component packages, \texttt{forestecology} is designed with verb-named
functions that can be modularly composed in sequence using the pipe
\texttt{\%\textgreater{}\%} operator \citep{bache_pipe_2020}. As we
articulate in Section \ref{casestudy}, these functions delineate the key
steps in our analysis sequence. Furthermore, the inputs and outputs of
nearly all of our functions use the same ``simple features for R'' data
structures as implemented in the \texttt{sf} package for standardized
support for spatial vector data \citep{pebesma_simple_2018}. The
\texttt{sf} package is a \texttt{tidyverse}-friendly evolution of the
\texttt{sp} package of classes and methods for spatial data in R
\citep{pebesma_sp_2005}. As such, wrangling and visualization spatial
data such as ours becomes much easier.

@@

We present a case-study of the \texttt{forestecology} package's use on
data from the Smithsonian Conservation Biology Institute's (SCBI) large
forest dynamics plot in Front Royal, Virginia, USA in Section
\ref{casestudy}, which is part of the ForestGEO global network of
research sites
\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}.
The package is designed with ForestGEO plot data in mind, but we
envision that it could easily be modified to work with data from other
forest plots, e.g.~the US Forest Service Forest Inventory and Analysis
plots or more generally to model interactions of any community of mapped
sessile organisms \citep{smith_forest_2002}.

@@

We demonstrate the \texttt{forestecology} package's functionality on
data from the Smithsonian Conservation Biology Institute (SCBI) large
forest dynamics plot, located at the Smithsonian's National Zoo and
Conservation Biology Institute in Front Royal, VA, USA
\citep{bourg_initial_2013}. The 25.6 ha (640 x 400 m) plot is located at
the intersection of three of the major physiographic provinces of the
eastern US---the Blue Ridge, Ridge and Valley, and Piedmont
provinces---and is adjacent to the northern end of Shenandoah National
Park.

@@

The output \texttt{growth\_scbi} is a single data frame of class
\texttt{sf} that includes variables \texttt{growth}, the average annual
growth in DBH (cm \(\cdot\) y\textsuperscript{-1}) for all stems that
were alive at both time points, and \texttt{geometry}, the \texttt{sf}
package's encoding of geolocations of type
\texttt{\textless{}POINT\textgreater{}}. In addition the species
variable \texttt{sp} is returned as a factor.\footnote{In our spatial
  cross-validation algorithm in Section \ref{spatial-cross-validation}
  issues can occur when rare species do not occur in the training set,
  but then are encountered in the test set. This risk is mitigated by
  representing \texttt{sp} as a factor variable, which has a complete
  list of all levels of the categorical variable.}

Given that \texttt{growth\_scbi} is of class \texttt{sf}, it can be
easily plotted in \texttt{ggplot2} using the \texttt{geom\_sf()}
geometry as seen in Figure \ref{fig:scbi-trees} where we plot a random
sample of 500 out of the 7954 trees.

@@

In the example below, we first manually define four folds that partition
the study region as an \texttt{sf} object. We then use the output of the
\texttt{spatialBlock()} function from the \texttt{blockCV} package to
associate each tree in \texttt{growth\_scbi} to the correct fold
\texttt{foldID} \citep{valavi_blockcv_2019}. \footnote{In the Supporting
  Information we present an example where the folds themselves are also
  created automatically using \texttt{spatialBlock()} given a specified
  \texttt{cv\_block\_size}, as opposed to manually as in the example.}

@@

Figure \ref{fig:scbi-spatial-information} illustrates the net effect of
adding these two elements of spatial information to
\texttt{growth\_scbi}. The location of each tree is marked with an
integer indicating its fold, where the folds are marked with solid
lines. The color of each digit indicates whether the tree is part of the
buffer region (and thus will only be considered as a competitor tree in
our model) or is part of the interior of the study region (and thus is a
focal tree whose growth is of modeled interest).

@@

\caption{Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4). All trees in the interior of the study region (i.e. not part of buffer) will be the focal trees whose growth will be modeled.}\label{fig:scbi-spatial-information}
\end{figure}

@@

The next step is to identify all focal trees and their corresponding
competitor trees. More specifically, identify all trees that are not
part of the buffer region, have a valid \texttt{growth} measurement, and
have at least one neighbor within 7.5m.
\texttt{create\_focal\_vs\_comp()} returns a new data frame of type
\texttt{sf}. On top of previously detailed arguments \texttt{comp\_dist}
and \texttt{id}, \texttt{create\_focal\_vs\_comp()} also requires an
\texttt{sf} object representation of the spatial cross-validation
blocks/folds as seen in Section \ref{spatial-information}. We present
the resulting data frame below with the \texttt{foldID} variable omitted
for compactness.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{foldID)}
\CommentTok{## # A tibble: 6,296 x 6}
\CommentTok{##   focal_ID focal_sp   dbh   geometry growth comp             }
\CommentTok{##      <dbl> <fct>    <dbl>    <POINT>  <dbl> <list>           }
\CommentTok{## 1        4 nysy     13.6  (14.2 428)  0.103 <tibble [20 x 4]>}
\CommentTok{## 2        5 havi      8.8   (9.4 436)  0.150 <tibble [32 x 4]>}
\CommentTok{## 3       79 tiam     47.7    (40 381) -0.161 <tibble [20 x 4]>}
\CommentTok{## 4       80 caca      5.15 (38.7 422)  0.253 <tibble [12 x 4]>}
\CommentTok{## 5       96 libe      2.3    (60 310)  0.262 <tibble [14 x 4]>}
\CommentTok{## # ... with 6,291 more rows}
\end{Highlighting}
\end{Shaded}

@@

We observe that for this focal tree, we have 4 variables of information
on its 20 competitor trees: their unique tree-stem ID number, their
distance to the focal tree (all \(\leq\) 7.5), their species, and their
basal area (in m\(^2\)) calculated as
\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) where \(DBH\) is the value
from the earlier of the two censuses in cm. Saving our focal versus
competitor information in list-column minimizes redundancy since we do
not repeat information on the focal tree 20 times. The spatial
distribution of these trees is visualized in Figure
\ref{fig:scbi-focal-vs-comp-map}: the dashed circle extends 7.5 m away
from the focal tree while all 20 competitor trees are within this
circle.

@@

\caption{Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4.}\label{fig:scbi-focal-vs-comp-map}
\end{figure}

@@

Lastly, the generic for \texttt{ggplot2::autoplot()} allows us to plot
the posterior distribution of all parameters in Figure
\ref{fig:scbi-posterior-viz} (for compactness we include only 3
species).

@@

\caption{Fit model: Posterior distributions of all parameters for three species.}\label{fig:scbi-posterior-viz}
\end{figure}

@@

To this end, we use the \texttt{foldID} variable defined in Section
\ref{spatial-information} whereby all focal trees are assigned to one of
4 spatially contiguous blocks that act as folds in our cross-validation
routine. Figure \ref{fig:scbi-spatial-cross-validation-schematic}
presents a schematic illustrating this scheme for fold 1 (bottom-left)
as the test set and folds 2--4 as the training set. We fit the model to
all focal trees in the training set, apply the model to all focal trees
in the test set to compute fitted/predicted values, and compute the RMSE
of the observed versus predicted growths. We repeat this procedure 3
more times with each of the three remaining folds acting as the test set
and then average all four resulting RMSE's. Furthermore, in order to
maintain spatial independence between the test and training set, a
``fold buffer'' that extend outwards from the boundary of the test set
is computed; all trees falling within this fold buffer are excluded from
the training set.

@@

\caption{Schematic of spatial cross-validation: Using the k = 1 fold as the test set, assigning each focal tree to training set, test set, and fold buffer.}\label{fig:scbi-spatial-cross-validation-schematic}
\end{figure}

@@

The \texttt{forestecology} package provides an accessible way to fit and
test models of neighborhood competition. Currently it is written to work
with data from ForestGEO plots, but it could easily be modified to work
on any single large, mapped forest plot in which at least two
measurements of each individual have been taken. With some data
wrangling the package could also be applied to forest inventory data in
which several small plots are spatially separated, e.g., USFS Forest
Inventory. In future versions of \texttt{forestecology} we also hope to
make it possible to model plant mortality in addition to plant growth.
The package follows the guidelines for \texttt{tidy} data, leverages the
\texttt{sf} package for spatial data, and S3 open-oriented model
structure. We hope that the package will increase the use of
neighborhood competition models to better understand what structures
plant competition.

@@

We intend to archive all data and source code for the
\texttt{forestecology} package as well as this manuscript on GitHub at
\url{https://github.com/rudeboybert/forestecology}. This repository will
be versioned and archived on Zenodo upon acceptance. The 2008 and 2014
Smithsonian Conservation Biology Institute census data loaded in Section
\ref{compute-growth}are available on GitHub at
\url{https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files}
and are versioned and archived on Zenodo at
\url{https://doi.org/10.5281/zenodo.2649301}
\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}.

>>>>

The package designed with ``tidy'' design principles in mind
\citep{wickham_welcome_2019}. Much like many of the \texttt{tidyverse}
component packages, \texttt{forestecology} is designed with verb-named
functions that can be modularly composed in sequence using the pipe
\texttt{\%\textgreater{}\%} operator \citep{bache_pipe_2020}. As we
articulate in Section \ref{casestudy}, these functions delineate the key
steps in our analysis sequence. Furthermore, the inputs and outputs of
nearly all of our functions use the same ``simple features for R'' data
structures as implemented in the \texttt{sf} package for standardized
and \texttt{tidyverse}-friendly support for spatial vector data
\citep{pebesma_simple_2018}

@@

@@

We present a case-study of the \texttt{forestecology} package's
functionality on data from the Smithsonian Conservation Biology
Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA,
which is part of the ForestGEO global network of research sites
\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}
\citep{bourg_initial_2013}. The 25.6 ha (640 x 400 m) plot is located at
the intersection of three of the major physiographic provinces of the
eastern US---the Blue Ridge, Ridge and Valley, and Piedmont
provinces---and is adjacent to the northern end of Shenandoah National
Park.

@@

The output \texttt{growth\_scbi} is a single data frame of class
\texttt{sf} that includes variables \texttt{growth}, the average annual
growth in DBH (cm \(\cdot\) y\textsuperscript{-1}) for all stems that
were alive at both time points, and \texttt{geometry}, the \texttt{sf}
package's encoding of geolocations of type
\texttt{\textless{}POINT\textgreater{}}. In addition the species
variable \texttt{sp} is returned as a factor.\footnote{In our spatial
  cross-validation algorithm in Section \ref{spatial-cross-validation}
  issues can occur when rare species do not occur in the training set,
  but then are encountered in the test set. This risk is mitigated by
  representing \texttt{sp} as a factor variable, which has a complete
  list of all levels of the categorical variable.} Given that
\texttt{growth\_scbi} is of class \texttt{sf}, it can be easily plotted
in \texttt{ggplot2} using the \texttt{geom\_sf()} geometry as seen in
Figure \ref{fig:scbi-trees}.

@@

In the example below, we first manually define four folds that partition
the study region as an \texttt{sf} object. We then use the output of the
\texttt{spatialBlock()} function from the \texttt{blockCV} package to
associate each tree in \texttt{growth\_scbi} to the correct fold
\texttt{foldID} \citep{valavi_blockcv_2019}. \footnote{In the Supporting
  Information we present an example where the folds themselves are also
  created automatically using \texttt{spatialBlock()} given a specified
  \texttt{cv\_block\_size}, as opposed to manually as in the example.}
Figure \ref{fig:scbi-spatial-information} illustrates the net effect of
adding these two elements of spatial information to
\texttt{growth\_scbi}.

@@

@@

\caption{Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4). The location of each tree is marked with an integer indicating its fold, with folds delineated with solid lines. The color of each digit indicates whether the tree is part of the buffer region (and thus will only be considered as a competitor tree in our model) or is part of the interior of the study region (and thus is a focal tree whose growth is of modeled interest).}\label{fig:scbi-spatial-information}
\end{figure}

@@

The next step is to identify all focal trees and their corresponding
competitor trees. More specifically, identify all trees that are not
part of the buffer region, have a valid \texttt{growth} measurement, and
have at least one neighbor within 7.5m.
\texttt{create\_focal\_vs\_comp()} returns a new data frame of type
\texttt{sf}. On top of previously detailed arguments \texttt{comp\_dist}
and \texttt{id}, \texttt{create\_focal\_vs\_comp()} also requires an
\texttt{sf} object representation of the spatial cross-validation
blocks/folds as seen in Section \ref{spatial-information}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, geometry, growth, comp)}
\CommentTok{## # A tibble: 6,296 x 5}
\CommentTok{##   focal_ID focal_sp   geometry growth comp             }
\CommentTok{##      <dbl> <fct>       <POINT>  <dbl> <list>           }
\CommentTok{## 1        4 nysy     (14.2 428)  0.103 <tibble [20 x 4]>}
\CommentTok{## 2        5 havi      (9.4 436)  0.150 <tibble [32 x 4]>}
\CommentTok{## 3       79 tiam       (40 381) -0.161 <tibble [20 x 4]>}
\CommentTok{## 4       80 caca     (38.7 422)  0.253 <tibble [12 x 4]>}
\CommentTok{## 5       96 libe       (60 310)  0.262 <tibble [14 x 4]>}
\CommentTok{## # ... with 6,291 more rows}
\end{Highlighting}
\end{Shaded}

@@

We observe that for this focal tree, we have 4 variables of information
on its 20 competitor trees: their unique tree-stem ID number, their
distance to the focal tree (all \(\leq\) 7.5), their species, and their
basal area (in m\(^2\)) calculated as
\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) where \(DBH\) is the value
from the earlier of the two censuses in cm. Saving our focal versus
competitor information in list-column minimizes redundancy since we do
not repeat information on the focal tree 20 times. The spatial
distribution of these trees is visualized in Figure
\ref{fig:scbi-focal-vs-comp-map}.

@@

\caption{Identify all focal and corresponding competitor trees: The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle.}\label{fig:scbi-focal-vs-comp-map}
\end{figure}

@@

Lastly, the generic for \texttt{ggplot2::autoplot()} allows us to plot
the posterior distribution of all parameters in Figure
\ref{fig:scbi-posterior-viz}.

@@

\caption{Fit model: Posterior distributions of all parameters. For compactness we include only three species.}\label{fig:scbi-posterior-viz}
\end{figure}

@@

To this end, we use the \texttt{foldID} variable defined in Section
\ref{spatial-information} whereby all focal trees are assigned to one of
4 spatially contiguous blocks that act as folds in our cross-validation
routine. Figure \ref{fig:scbi-spatial-cross-validation-schematic}
presents a schematic illustrating this scheme. We fit the model to all
focal trees in the training set, apply the model to all focal trees in
the test set to compute fitted/predicted values, and compute the RMSE of
the observed versus predicted growths. We repeat this procedure 3 more
times with each of the three remaining folds acting as the test set and
then average all four resulting RMSE's. Furthermore, in order to
maintain spatial independence between the test and training set, a
``fold buffer'' that extend outwards from the boundary of the test set
is computed; all trees falling within this fold buffer are excluded from
the training set.

@@

\caption{Schematic of spatial cross-validation: Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a "fold buffer."}\label{fig:scbi-spatial-cross-validation-schematic}
\end{figure}

@@

The \texttt{forestecology} package provides an accessible way to fit and
test models of neighborhood competition. While the package is designed
with ForestGEO plot data in mind, we envision that it can be modified to
work on i) any single large, mapped forest plot in which at least two
measurements of each individual have been taken, e.g.~the US Forest
Service Forest Inventory and Analysis plots or ii) more generally to
model interactions of any community of mapped sessile organisms
\citep{smith_forest_2002}. In future versions of \texttt{forestecology}
we also hope to make it possible to model plant mortality in addition to
plant growth. The package follows the guidelines for \texttt{tidy} data,
leverages the \texttt{sf} package for spatial data, and S3 open-oriented
model structure. We hope that the package will increase the use of
neighborhood competition models to better understand what structures
plant competition.

@@

We intend to archive all data and source code for this manuscript on
GitHub at \url{https://github.com/rudeboybert/forestecology}. This
repository will be archived on Zenodo upon acceptance. The example
Smithsonian Conservation Biology Institute census data used are
available on GitHub at
\url{https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files}
and are archived on Zenodo at
\url{https://doi.org/10.5281/zenodo.2649301}
\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}.
